EDIpack 2.0: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-20 05_examples.tex
Iterated examples parts. Ready to share.
* To be iterated Intro and Conclusions (that is on me)
* Missing discussion of the conjugate gradient fit of matrix GF norms (on Gabriel)
* Totally incomplete bibliograpy.
* Missing finite Temperature comparison between EDIpack and W2Dyn (on Alexander+Lorenzo?)

\subsection{Bethe lattice DMFT}\label{SecExamplesBetheDMFT}
The test bed for any DMFT application is conventionally considered 
the description of the Mott or metal-to-insulator transition (MIT) 
in a Bethe lattice, i.e., a Cayley tree with coordination number 
$z\to\infty$~\cite{Georges1996}.
Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw}
    $$

@@

\subsubsection{W2Dynamics solution}\label{SecExamplesBetheDMFTw2d}
To demonstrate use of the w2dynamics interface, here we briefly
discuss how to solve the DMFT problem for the Bethe lattice using
W2Dynamics interface. \fixme{For definitiness we compare the finite low
temperature results obtained via the two solvers: ED and CTQMC.  }

@@

\subsection{Attractive Hubbard model}\label{SecExamplesAHM}
The second example we present concerns the DMFT description of the 
attractive Hubbard model on a two-dimensional square lattice. This 
example has a twofold aim: (i) to demonstrate the support of \NAME 
for ($s$-wave) superconductivity, and (ii) to illustrate the use of 
the library's Python API through a concrete working example. The 
model Hamiltonian is given by:
$$
H = \sum_{\ka,\s} \epsilon(\ka) c^\dagger_{\ka\s} c_{\ka\s} 
    - U \sum_i n_{i\up} n_{i\dw},
$$
where $U > 0$, $c^\dagger_{\ka\s} = \tfrac{1}{\sqrt{N}} 
\sum_i e^{-i\ka \cdot R_i} c^\dagger_{i\s}$, and $c^\dagger_{i\s}$ 
is the creation operator for an electron at site $i$ with spin $\s$. 
The occupation operator is $n_{i\s} = c^\dagger_{i\s} c_{i\s}$, and 
the energy dispersion relation is $\e(\ka)=-2t[\cos{(k_x a)}+
\cos{(k_y a)}]$, where we set the lattice spacing 
$a=1$  and the choose the energy unit such as $4t=D=1$ for convenience.

@@

\subsection{Multi-orbital impurity with Hubbard-Kanamori interaction}
\label{SecExamplesTRIQS}

@@

\subsection{Interacting Bernevig-Hughes-Zhang model}
In this section we focus on the DMFT solution of the interacting
Bernevig-Hughes-Zhang (BHZ).
As shown in Ref.~\onlinecite{AmaricciPRB2022}, at moderate
interactions this model admits an excitonic phase in which opposite
orbitals electrons and holes across the band gap bound into a coherent phase.

@@

In the following we consider the case of $M=1$ which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$. 
%
The main effect of the interaction is enclosed in the renormalized
mass term $M_{eff}=M+\tfrac{1}{4}\Tr{\Re\Sigma(i\omega\to0)}$. The
real-part of the self-energy being proportional to $\Gamma_{03}$
corrects the mass term with respect to its bare value. To leading
order this correction is given proportional to $\tfrac{U-5J}{4}\langle
\hat{T}_z\rangle$. So, for our choice of parameters, the effect of
interaction would be to reduce the (effective) mass term and thus
reducing the topological band gap. In the strong-coupling limit the
two orbitals get populated with one electron per site, reaching the
conditions for the formation of a Mott insulating state.
As reported in panel (A) of \figu{figEx4}, at intermediate coupling,
the DMFT solutions features the formation of a
region of exciton condensation with $\langle E_x\rangle>0$, $\langle
E_0\rangle=\langle E_z\rangle=0$. 
Unlike the static mean-field description~\cite{Blason2021}, the
transition from the QSHI topological state to the Excitonic Insulator 
(EI) becomes of first-order in presence of quantum fluctuations
corrections~\cite{Continentino,Paoletti} contained in DMFT. On the
other hand, the EI continuously evolves into a Mott Insulator (MI)
for larger interaction (neglecting for simplicity any 
anti-ferromagnetic ordering that would naturally occurr in this
regime).
In the same panel we report the evolution of 
$\langle T_z\rangle$ to illustrate the progressive reduction of the
orbital polarization and, thus, of the effective mass.

@@

\subsection{Holstein model on the Bethe lattice (Fortran)}

One of the features introduced in \NAME is the possibility of treating local phonons in the superconductive case (i.e. \textbf{ed\_mode=superc} ). In this example we show how to get the normal and superconductive solutions of the Holstein model on the Bethe lattice within DMFT. The phononic and electron-phonon part of the Holstein Hamiltonian at half-filling for particle-hole symmetric density of states like in the Bethe lattice reads:
\begin{equation} \label{eqex:H_Holstein}
    H_{int} = \sum_i \Big[\omega_0 b^\dagger_i b_i + g(b^\dagger_i + b_i)\sum_{\sigma}(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2})\Big]
\end{equation}

In the following we show for a fixed value of $\lambda/D=1.0$ the adiabatic to anti-adiabatic crossover for the uniform solution of the Holstein problem on the Bethe lattice.

The Holstein interaction takes a simpler form for $\omega_0 \rightarrow \infty$ with $g^2/\omega_0=\lambda/2$ constant, also known as anti-adiabatic limit, for which it becomes a Hubbard attraction:
\begin{equation}
    H_{int} \overset{ \omega_0 \rightarrow \infty}{ \longrightarrow } -\frac{\lambda}{2} \sum_i \Big[\sum_\sigma(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2}) \Big]^2
\end{equation}

Meanwhile, in the opposite limit, $\omega_0 \rightarrow 0$ with $g^2/\omega_0=\lambda/2$ constant, also known as adiabatic limit, the system is in a phase of Bipolaronic insulator~\citep{Capone2006PRB} for $\lambda=1$.

\begin{figure}[ht!]
    \includegraphics[width=\linewidth]{edipack2_examples/Bethe_Holstein/figBethe_Holstein.pdf}
    \caption{\label{figEx5}Plots of the double occupancy (A) and the superconductive order parameter (B) as a function of $\omega_0/D$ for $\lambda/D=1$ in the normal (orange) and superconductive (red) phase. The horizontal broken line are the values for the corresponding Hubbard model in the anti-adiabatic limit.}
\end{figure}

>>>>

\subsection{Hubbard model on the Bethe lattice (Fortran API)}\label{SecExamplesBetheDMFT}
The test bed for any DMFT application is conventionally considered 
the description of the Mott or metal-to-insulator transition (MIT) 
in a Bethe lattice, i.e., a Cayley tree with coordination number 
$z\to\infty$~\cite{Georges1996}.
Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw}
    $$

@@

\subsubsection{Finite temperature solution (W2Dynamics interface)}\label{SecExamplesBetheDMFTw2d}
To demonstrate use of the w2dynamics interface, here we briefly
discuss how to solve the DMFT problem for the Bethe lattice using
W2Dynamics interface. \fixme{For definitiness we compare the finite low
temperature results obtained via the two solvers: ED and CTQMC.  }

@@

\subsection{Attractive Hubbard model (Python API, {\tt ed\_mode=superc})}\label{SecExamplesAHM}
The second example we present concerns the DMFT description of the 
attractive Hubbard model on a two-dimensional square lattice. This 
example has a twofold aim: (i) to demonstrate the support of \NAME 
for ($s$-wave) superconductivity, and (ii) to illustrate the use of 
the library's Python API through a concrete working example. The 
model Hamiltonian is given by:
$$
H = \sum_{\ka,\s} \epsilon(\ka) c^\dagger_{\ka\s} c_{\ka\s} 
    - U \sum_i n_{i\up} n_{i\dw},
$$
where $U > 0$, $c^\dagger_{\ka\s} = \tfrac{1}{\sqrt{N}} 
\sum_i e^{-i\ka \cdot R_i} c^\dagger_{i\s}$, and $c^\dagger_{i\s}$ 
is the creation operator for an electron at site $i$ with spin $\s$. 
The occupation operator is $n_{i\s} = c^\dagger_{i\s} c_{i\s}$, and 
the energy dispersion relation is $\e(\ka)=-2t[\cos{(k_x a)}+
\cos{(k_y a)}]$, where we set the lattice spacing 
$a=1$  and the choose the energy unit such as $4t=D=1$ for convenience.

@@

\subsection{Multi-orbital quantum impurity with Hubbard-Kanamori
  interaction (TRIQS interface)}
\label{SecExamplesTRIQS}

@@

\subsection{Interacting Bernevig-Hughes-Zhang model (Fortran API, {\tt ed\_mode=nonsu2})}
In this section we focus on the DMFT solution of the interacting
Bernevig-Hughes-Zhang (BHZ).
As shown in Ref.~\onlinecite{AmaricciPRB2022}, at moderate
interactions this model admits an excitonic phase in which opposite
orbitals electrons and holes across the band gap bound into a coherent phase.

@@

In the following we consider the case of $M=1$ which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$.
The implemention is nearly identical to the cases discussed above and
we report it here for completeness:
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
     iloop=iloop+1
     !> Solve the impurity problem, retrieve $\Sigma(i\omega)$
     call ed_solve(bath)
     call ed_get_sigma(Smats,axis='mats')
     !> Get $G_{loc}$ (assume to have a function for this task)
     call get_gloc(Hk,Gmats,Smats,axis='m')
     !> Update the Weiss field $\GG^{-1}=G^{-1}_{loc}+\Sigma$ +  Linear mixing.
     call dmft_self_consistency(Gmats,Smats,Weiss)
     if(iloop>1)Weiss = wmixing*Weiss + (1.d0-wmixing)*Weiss_;Weiss_=Weiss
     !> Fit to updated the bath
     call ed_chi2_fitgf(Weiss,bath,ispin=1)
     !Check convergence (using DMFT_TOOLS)
     converged = check_convergence(Weiss(1,1,:),dmft_error,nsuccess,nloop)
   enddo  
\end{lstlisting}

%
The main effect of the interaction is enclosed in the renormalized
mass term $M_{eff}=M+\tfrac{1}{4}\Tr{\Re\Sigma(i\omega\to0)}$. The
real-part of the self-energy being proportional to $\Gamma_{03}$
corrects the mass term with respect to its bare value. To leading
order this correction is given proportional to $\tfrac{U-5J}{4}\langle
\hat{T}_z\rangle$. So, for our choice of parameters, the effect of
interaction would be to reduce the (effective) mass term and thus
reducing the topological band gap. In the strong-coupling limit the
two orbitals get populated with one electron per site, reaching the
conditions for the formation of a Mott insulating state.
As reported in panel (A) of \figu{figEx4}, at intermediate coupling,
the DMFT solutions features the formation of a
region of exciton condensation with $\langle E_x\rangle>0$, $\langle
E_0\rangle=\langle E_z\rangle=0$. 
Unlike the static mean-field description~\cite{Blason2021}, the
transition from the QSHI topological state to the Excitonic Insulator 
(EI) becomes of first-order in presence of quantum fluctuations
corrections~\cite{Continentino,Paoletti} contained in DMFT. On the
other hand, the EI continuously evolves into a Mott Insulator (MI)
for larger interaction (neglecting for simplicity any 
anti-ferromagnetic ordering that would naturally occurr in this
regime).
In the same panel we report the evolution of 
$\langle T_z\rangle$ to illustrate the progressive reduction of the
orbital polarization and, thus, of the effective mass.

@@

\subsection{Holstein model on the Bethe lattice (Fortran API, {\tt electron-phonons})}
One of the features introduced in \NAME is the possibility of treating
local phonons in conjunction with the superconductive phase,
i.e. \textbf{ed\_mode=superc}. 
In this example discuss  the normal and
superconductive solution of the pure Holstein model on the Bethe
lattice within DMFT.
We stress that, using Fortran API, the code
implementation for this specific case is essentially identical to the listing reported in the sections
\ref{SecExamplesBetheDMFT} ({\tt ed\_mode=normal}) and
\ref{SecExamplesAHM} ({\tt ed\_mode=superc}).

So we consider the model introduced in \secu{SecExamplesBetheDMFT}
with $U=0$ and the additional phononic and electron-phonon terms:
\begin{equation} \label{eqex:H_Holstein}
    H_{int} = \sum_i \Big[\omega_0 b^\dagger_i b_i + g(b^\dagger_i +
    b_i)\sum_{\sigma}(c^\dagger_{i\sigma}c_{i\sigma}
    -\frac{1}{2})\Big]. 
\end{equation}
We consider the half-filling regime of the particle-hole symmetric
Bethe lattice density of states. We set the half-bandwidth as our
energy unit $D=1$ and introduced the electron-phonon coupling $\lambda
= \tfrac{2g^2}{\omega_0D}$.  
The iterative DMFT solution algorithm follows the same principles
illustrated in the previous sections.

\paragraph{Results}
In the following we discuss the adiabatic ($\omega_0\to0$) to
anti-adiabatic ($\omega_0\to\infty$) crossover for
the uniform solution of the Holstein model at constant coupling $\lambda=1.0$.

In the anti-adiabatic limit, the Holstein interaction takes a
particularly simple form:
\begin{equation}\label{HlikeAttraction}
    H_{int} \overset{ \omega_0 \rightarrow \infty}{ \longrightarrow } -\frac{\lambda}{2} \sum_i \Big[\sum_\sigma(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2}) \Big]^2
\end{equation}
which describes a local Hubbard-like attraction among electrons
(mediated by local phonons).
In the opposite, adiabatic, limit, $\omega_0 \rightarrow 0$ the system
enters in a Bipolaronic Insulating phase~\citep{Capone2006PRB} for our
choice of the coupling.

We characterize the model solution by showin the evolution of the
double occupation $\langle n_{\up}n_{\dw}\rangle$ as a function of the
phonon frequency $\omega_0$, see panel (A). In this panel we compare
the behavior for {\tt ed\_mode=normal} against the superconductive
solution ({\tt ed\_mode=superc}). Our results capture the whole
crossover from adiabatic to anti-adiabatic regime. In the former
regime, the double occupation takes a similar value for the two
phases. However, approaching the anti-adiabatic regime, the two
solutions reach the limiting values corresponding to the residual
attraction \equ{HlikeAttraction}.   
To better characterize the nature of the superconducting
phase in the Holstein model we report the evolution of the anomalous order
parameter $\phi$ in the adiabatic to anti-adiabatc crossover.
The DMFT results obtained with \NAME show the rapid increase of the
superconducting order parameter as phonon frequency grows
large. Finally, in the anti-adiabatic regime $\phi$ saturates to a
finite value corresponding to the attractive Hubbard-like interaction
of strenght $\lambda/2$.

\begin{figure}[ht!]
    \includegraphics[width=\linewidth]{figures/figBethe_Holstein.pdf}
    \caption{\label{figEx5}
      Evolution of the  double occupation (A) and superconductive order parameter (B) as a function of $\omega_0/D$ for $\lambda=1$ in the normal (orange) and superconductive (red) phase. The horizontal broken line are the values for the corresponding Hubbard model in the anti-adiabatic limit.}
\end{figure}
